---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# `JuliaSwitch`

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/patter)](https://CRAN.R-project.org/package=patter)

`JuliaSwitch` is a high-level `R`--`Julia` interface. The package provides a common syntax that enables users to switch between multiple backends for `R`--`Julia` communication without changes to the code base.

* **Run `Julia` code** via a `julia_*()` helper function or `julia_cmd()`;
* **Push objects to `Julia`** using `julia_push()`;
* **Pull objects back** via `julia_pull()`;

# Motivation

`JuliaSwitch` was motivated our work on animal tracking with `patter`. This package provides an `R` interface to `Patter.jl`, which fit state-space models to animal-tracking data in `Julia`. 

> When `patter` was first developed, we used [`JuliaCall`](https://github.com/JuliaInterop/JuliaCall) as a `R`--`Julia` interface. That package connects `R` and `Julia` via a `C` interface, which is fast but can be unstable across `Julia` versions and platforms. (We experienced particular challenges on Linux when using `R` and `Julia` packages that require the same dynamic libraries.) 

> [`JuliaConnectoR`](https://github.com/stefan-m-lenz/JuliaConnectoR) is another `R`--`Julia` option.  Unlike `JuliaCall`, this package interfaces `R` and `Julia` using Transmission Control Protocol. This is a looser form of integration which trades speed for stability. 

`JuliaSwitch` became the solution to toggle between these two options. 

# Installation

`JuliaSwitch` can be installed with: 

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("edwardlavender/JuliaSwitch", 
                         dependencies = TRUE)
```

# Functionality 

* Run `julia_backend("JuliaCall")` or `julia_backend("JuliaConnectoR")` to set the `Julia` backend. 

* Start a `Julia` session via `julia_start()`. 

* Run arbitrary `Julia` code via `julia_cmd()`, `julia_cmd_line()` and `julia_cmd_block()`.

* For selected operations, helper functions are provided:
    - `julia_pkg_activate()` activates a local environment;
    - `julia_pkg_add()` and `julia_pkg_update()` add/update package(s);
    - `julia_pkg_installed()` checks if package(s) are installed;
    - `julia_using()` and `julia_import()` load and import package(s);
    - `julia_include()` sources a `Julia` script;
    - `julia_defined()` checks if an object is defined;
    - `julia_println()` prints lines;
    - `julia_helpfile()` prints a help file;
    - `julia_save()` and `julia_load()` save and load a file;

* To push an object from `R` to `Julia`, use `julia_push()`.

* To pull an object back from `Julia` to `R`, use `julia_pull()`. 

* Stop a `Julia` session via `julia_stop()`. 

# `S3` methods

`JuliaSwitch` uses `S3` methods to handle object transfers between `R` and `Julia`. Method dispatch is implemented by the internal functions `julia_send()`, `juliaSend()`, `julia_receiver()` and  `juliaReceive()`.The default methods call the relevant `JuliaCall` or `JuliaConnectoR` routines. Special cases (such as `data.frame`s and `terra::SpatRaster`s) are handled by custom methods. Specify additional methods for other special cases. 

# Citation

**To cite `JuliaSwitch` in publications**, please use:

Lavender, E., Albert, C., & Scheidegger, A. (2025). Animal geolocation with convolution algorithms in Julia and R via Wahoo.jl. Methods in Ecology and Evolution, 00, 1–8. https://doi.org/10.1111/2041-210x.70185

```bibtex
@article{Lavender2025Wahoo,
  author  = {Lavender, Edward and Albert, Carlo and Scheidegger, Andreas},
  title   = {Animal geolocation with convolution algorithms in Julia and R via Wahoo.jl},
  journal = {Methods in Ecology and Evolution},
  doi     = {10.1111/2041-210X.70185},
  url     = {https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.70185},
  keywords = {biologging, biotelemetry, filter, hidden Markov model, movement ecology, package, passive acoustic telemetry, state-space model}
}
```

# Code of conduct

Please note that the `JuliaSwitch` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
